Using Deep Learning to Identify Patients with Cognitive Impairment in Electronic Health Records

### 2021-09-12 Submission/Paper/ML4H/Paper.tex
first round of citations

\section{Related Works}
\label{sec:RelatedWorks} 
Prior works have used NLP techniques to detect various diseases from EHR. Rajkmoar et al. % https://www.nature.com/articles/s41746-018-0029-1%22
used recurrent neural networks (long short-term memory (LSTM)) among others to predict inpatient mortality using EHR data from the University of California, San Francisco (UCSF) from 2012 to 2016, and the University of Chicago Medicine (UCM) from 2009 to 2016.  Glicksberg et al. % https://www.worldscientific.com/doi/epdf/10.1142/9789813235533_0014 
performed phenotyping for diseases such as Attention Deficit Hyperactivity Disorder (ADHD) by clustering on word2vec embeddings from EHR of the Mount Sinai Hospital (MSH) in New York City. These studies have shown that the application of NLP techniques to EHR have improved disease detection, and that NLP techniques can be applied to dementia detecton to achieve similar results. Our work uses deep learning NLP techniques, which has achieved impressive results when applied to general text due to the use of word embeddings and attention-based models, but has had limited applications in healthcare. %(Vaswani et al., 2017; Mikolov et al., 2013; Pennington et al., 2014; Peters et al., 2018; Devlin et al., 2018),

@@

\label{sec:TFIDF}  
\paragraph{(1) Logistic Regression with TF-IDF Vectors} We performed TF-IDF (term frequency-inverse document frequency) vectorization on the annotated sequences and selected features based on a term's Pearson correlation coefficient with the outcome. L1 Regularized logistic regression was applied with the annotated cognitive impairment labels. We used 10-fold cross validation to select different correlation coefficients as thresholds to select features and iterated over different lambda values to determine the optimal lambda value and correlation coefficient threshold.

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called ClinicalBERT, which was trained on the MIMIC II database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformer and Simpletransformers packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Optuna was used to perform a 20-trial study and tune the learning rate, adam epilson, and the number of train epochs on the held-out validation set to maximize AUC. An early stopping rule of 3 epochs was also used to prevent overfitting.

>>>>

\section{Related Works}
\label{sec:RelatedWorks} 
Prior works have used NLP techniques to detect various diseases from EHR. \citep{rajkomar2018scalable} % https://www.nature.com/articles/s41746-018-0029-1%22
used recurrent neural networks (long short-term memory (LSTM)) among others to predict inpatient mortality using EHR data from the University of California, San Francisco (UCSF) from 2012 to 2016, and the University of Chicago Medicine (UCM) from 2009 to 2016.  \citep{glicksberg2018automated} % https://www.worldscientific.com/doi/epdf/10.1142/9789813235533_0014 
performed phenotyping for diseases such as Attention Deficit Hyperactivity Disorder (ADHD) by clustering on word2vec embeddings from EHR of the Mount Sinai Hospital (MSH) in New York City. These studies have shown that the application of NLP techniques to EHR have improved disease detection, and that NLP techniques can be applied to dementia detecton to achieve similar results. Our work uses deep learning NLP techniques, which has achieved impressive results when applied to general text due to the use of word embeddings and attention-based models \citep{vaswani2017attention,mikolov2013distributed,pennington2014glove,peters2018deep, devlin2018bert}, but have had limited applications in healthcare. %(Vaswani et al., 2017; Mikolov et al., 2013; Pennington et al., 2014; Peters et al., 2018; Devlin et al., 2018),

@@

\label{sec:TFIDF}  
\paragraph{(1) Logistic Regression with TF-IDF Vectors} We performed TF-IDF (term frequency-inverse document frequency) vectorization on the annotated sequences and selected features based on a term's Pearson correlation coefficient with the outcome. L1 Regularized logistic regression \citep{tibshirani1996regression} was applied with the annotated cognitive impairment labels. We used 10-fold cross validation to select different correlation coefficients as thresholds to select features and iterated over different lambda values to determine the optimal lambda value and correlation coefficient threshold.

@@

\label{sec:Transformer}  
\paragraph{(2) Transformer Based Sequence Classification Language Model} We utilized a pre-trained language model called ClinicalBERT \citep{alsentzer2019publicly}, which was trained on the MIMIC II database containing EHR records from ICU patients. We used the implementation in the Huggingface Transformer \citep{Wolf2019HuggingFacesTS} and Simpletransformers \citep{simple2020thilina} packages. After text preprocessing, input texts were tokenized with the default tokenizer and converted to embeddings. The model was initialized with pre-trained parameters and later fine-tuned on our labeled training set. Optuna \citep{akiba2019optuna} was used to perform a 20-trial study and tune the learning rate, adam epilson, and the number of train epochs on the held-out validation set to maximize AUC. An early stopping rule of 3 epochs was also used to prevent overfitting.
